---
title: "Assignment-1"
author: "Jiwon Park"
date: '2020 9 9 '
output: html_document
---
```{r message=FALSE} 
library(tidycensus)
library(tidyverse)
library(dplyr)
```

## Variables Used

```{r}
area_vars_2018 <- load_variables (2018, "acs5")
```

1. Total population (continuous) : tot_pop
2. Median Income (continuous) : med_income
3. Monthly Housing Cost (continuous) : mon_hous
4. Percentage of the tenant population (continuous) : pct_ten
5. Percentage of the foreign-born population (continuous) : pct_foreign
6. Majority Level of Education (categorical) : maj_edu
7. Majority Mode of Transportation (categorical) : maj_tra
8. Majority race of the population (not Hispanic or Latino) (categorical) : maj_race

```{r message=FALSE}
pop_vars <- c(tot_pop = "B01003_001",
              med_income = "B21004_001",
              mon_hous = "B25104_001", 
              tot_ten = "B07013_003", 
              tot_foreign = "B05002_013",
              tot_less_high = "B06009_002",
              tot_high = "B06009_003",
              tot_col = "B06009_004",
              tot_bac = "B06009_005",
              tot_gra = "B06009_006",
              tot_car = "B08006_002",
              tot_pubtra = "B08006_008",
              tot_bic = "B08006_014",
              tot_walk = "B08006_015",
              tot_tra_other = "B08006_016",
              nh_white_alone = "B03002_003",
              nh_black_AA_alone = "B03002_004",
              nh_asian_alone =  "B03002_006",
              h_latino = "B03002_012",
              nh_race_other = "B03002_008")

tractLA2018 <- get_acs(geography = "tract",
                        variables = pop_vars,
                        output = "wide",
                        state = "CA",
                       county = "Los Angeles",
                        year = 2018,
                        survey = "acs5")
```

## Calculating new variables

```{r}
tractLA2018 <- tractLA2018 %>%
  mutate(pct_ten = tot_tenE / tot_popE,
         pct_foreign = tot_foreignE / tot_popE,
         pct_less_high = tot_less_highE / tot_popE,
         pct_high = tot_highE / tot_popE,
         pct_col = tot_colE / tot_popE,
         pct_bac = tot_bacE / tot_popE,
         pct_gra = tot_graE / tot_popE,
         pct_car = tot_carE / tot_popE,
         pct_pubtra = tot_pubtraE / tot_popE,
         pct_bic = tot_bicE / tot_popE,
         pct_walk = tot_walkE / tot_popE,
         pct_tra_other = tot_tra_otherE / tot_popE,
         pct_nh_white = nh_white_aloneE / tot_popE,
         pct_nh_black = nh_black_AA_aloneE / tot_popE,
         pct_h_latino = h_latinoE / tot_popE,
         pct_nh_asian = nh_asian_aloneE / tot_popE,
         pct_nh_race_other = nh_race_otherE / tot_popE) %>%
  mutate(pct_less_col = pct_less_high + pct_less_high,
         pct_col = pct_col + pct_bac) %>%
  mutate(maj_edu = case_when(pct_less_col > 0.5 ~ "Less than College",
                             pct_col > 0.5 ~ "College and Above")) %>%
  mutate(maj_tra = case_when(pct_car > 0.5 ~ "Car",
                             pct_pubtra > 0.5 ~ "Public Transportation",
                             pct_bic > 0.5 ~ "Bicycle",
                             pct_walk > 0.5 ~ "Walk",
                             pct_tra_other > 0.5 ~ "Other",
                             TRUE ~ "No Majority")) %>%
  mutate(maj_race = case_when(pct_nh_white > 0.5 ~ "nh_white",
                             pct_nh_black > 0.5 ~ "nh_black",
                             pct_nh_asian > 0.5 ~ "nh_asian",
                             pct_h_latino > 0.5 ~ "hs_latino",
                             pct_nh_race_other > 0.5 ~ "nh_other",
                             TRUE ~ "no_majority")) %>%
  select(tot_popE, med_incomeE, mon_housE, pct_ten, pct_foreign, maj_edu, maj_tra, maj_race)

tractLA2018
```

## Saving the data

My final dataset has 2,346 rows, which mean my analysis will include 2,346 census tracts in the LA county.

Finally, I'll save my data as a CSV file.

```{r}
write_csv(tractLA2018, "tractLA2018")
```


